\documentclass{article}

\usepackage[submission]{../../../colm/2026/colm2026_conference}
\usepackage{array}
\usepackage{booktabs}
\usepackage{url}
\usepackage{xurl}
\usepackage{hyperref}

\title{Outlier Migration in Long Decode Traces}
\author{Anonymous authors\\Paper under double-blind review}

\begin{document}
\maketitle

\begin{abstract}
This draft reports an observational measurement candidate for OutlierMigrate,
studying whether high-magnitude activation channels remain stationary during
long decode. It is not a camera-ready positive systems method: Phase 0 passes on
Granite-4.0-H-Tiny: the fraction of top-1\% position-100 channels whose rank
changes by more than two channel positions by decode position 10000 is
0.8178385416666667 with bootstrap 95\% CI
[0.797265625, 0.8368489583333334]. Phase 1 replicates the dynamic-outlier
direction at larger scale on Granite-4.0-H-Small: the migration fraction is
0.843165650406504 with bootstrap 95\% CI
[0.8334349593495936, 0.8511432926829268]. A partial Phase 2 cross-family check
on \texttt{nvidia/NVIDIA-Nemotron-3-Nano-30B-A3B-BF16} also passes:
migration fraction 0.820809591642925 with bootstrap 95\% CI
[0.7865325261158594, 0.8544931149097815]. This is evidence that the
rank-migration surface is not limited to the measured Granite runs, but it is
not full cross-validation: Qwen3.6 and Kimi Linear are deferred pending runtime
compatibility. The missing migration-aware intervention and the deferred
Qwen3.6/Kimi runs are blockers for any positive-method or systems-efficiency
claim.
\end{abstract}

\section{Introduction}

OutlierMigrate is currently scoped as an observational measurement study: it
asks whether the channels carrying high-magnitude decode-time activations are
fixed or whether they move during long reasoning traces. Static
outliers would support static channel-protection recipes. Dynamic outliers
could motivate migration-aware protection or routing methods, but this draft
does not implement or validate such a method. That step requires intervention
evidence showing the effect is stable, useful, and not specific to one model
family.

The current paper status is deliberately limited. Phase 0 is a screen on
\texttt{ibm-granite/granite-4.0-h-tiny} with 12 deterministic AIME-2025
reasoning traces. Phase 1 scales the same dynamic path to
\texttt{ibm-granite/granite-4.0-h-small} with 24 deterministic traces and an
extended decode-position grid. A partial Phase 2 run applies the same
measurement to Nemotron-3-Nano-30B-A3B-BF16 under the authorized no-upgrade
window. Together they support replicated same-family measurement evidence and
one partial cross-family check under the preregistered metric. They do not
support a camera-ready final claim, a positive systems method, an end-to-end
efficiency claim, or a completed Qwen3.6/Kimi Linear validation claim.

\section{Related Work}

\paragraph{Mamba quantization and dynamic outliers.}
OutlierMigrate is not the first work to observe dynamic outlier behavior in
Mamba-style models. In vision state-space models, QMamba reports that hidden
state sequences exhibit highly dynamic variations and introduces Temporal
Group Quantization for hidden states \citep{li2025qmamba}. OuroMamba makes the
point even more directly for vision Mamba models: it identifies dynamic
activation-outlier variation across time steps as a failure mode for static
post-training quantization and uses lightweight dynamic outlier detection
during inference \citep{ramachandran2025ouromamba}. These results are important
prior art. They establish that dynamic outliers are already known in the
vision-Mamba setting.

Language-Mamba quantization work has emphasized a different operating regime.
Mamba-PTQ shows that recurrent LLMs exhibit activation outlier channels and
frames outlier-aware quantization as a first step \citep{pierro2024mambaptq}.
Quamba proposes a static 8-bit per-tensor recipe that suppresses maximum input
values to selective SSMs and uses a Hadamard transform for output activations
\citep{chiang2024quamba}. MambaQuant uses variance-aligned rotations and
reports heavy-tailed, channel-uneven distributions in Mamba-family models
\citep{xu2025mambaquant}. Quamba-SE softens the treatment of activation
outliers with multiple adaptive scales but still evaluates a per-value
quantizer rather than a decode-time channel migration policy
\citep{chen2026quambase}. OutlierMigrate asks which regime applies to hybrid
LLMs in long reasoning traces: the static or calibration-time protection
regime common in language deployment recipes, or the dynamic regime already
documented in vision Mamba.

\paragraph{Static protection in LLM deployment.}
Several influential LLM quantization systems protect or transform salient
activation structure using calibration-time statistics or fixed transformations.
SmoothQuant migrates activation difficulty into weights using offline
activation statistics \citep{xiao2023smoothquant}; AWQ protects a small set of
activation-salient weight channels \citep{lin2023awq}; QuaRot removes hidden
state outliers with rotations \citep{ashkboos2024quarot}; KVQuant isolates KV
cache outliers per vector and uses per-channel key quantization
\citep{hooper2024kvquant}; and BlockDialect assigns block-wise number formats
from a learned formatbook \citep{jang2025blockdialect}. These systems are not
claimed to be wrong by our current evidence. The implication is narrower: if
the protected high-magnitude set changes substantially during long decode in
hybrid LLMs, static protection maps may need migration-aware updates or
validation on the specific recurrent/linear-attention architecture. Several
of these systems already include adaptive components, such as KVQuant's
per-vector dense-and-sparse handling or BlockDialect's online activation
format selection; our current evidence motivates stress-testing those designs
on hybrid long decode, not declaring them failed.

\paragraph{Channel-wise gated recurrence.}
Recent efficient-recurrence and linear-attention architectures make
data-dependent channel-wise memory control a central design axis. Gated Linear
Attention (GLA) formulates linear attention as recurrent matrix-valued state
with data-dependent gating \citep{yang2023gla}. Gated DeltaNet combines
adaptive memory erasure with delta-rule updates and reports hybrids that mix
Gated DeltaNet with attention or Mamba2 layers \citep{yang2024gdn}. Moonshot
AI's October 2025 Kimi Linear report is a recent example in this line: its
Kimi Delta Attention (KDA) is described as channel-wise gated linear attention
that extends Gated DeltaNet with finer-grained gating, and the released Kimi
Linear model is a layerwise hybrid of KDA and Multi-Head Latent Attention
\citep{kimilinear2025}. The official Qwen3.6 model card likewise describes a
hybrid layout that interleaves Gated DeltaNet blocks with gated attention
\citep{qwen36hf}. These sources motivate OutlierMigrate as an empirical test
of whether high-magnitude channels stay fixed during long decode in hybrid
language models. They do not imply that Kimi Linear, Qwen3.6, GLA, or other
unmeasured models exhibit the migration measured in this draft.

\section{Method}

Phase 0 follows the frozen preregistration at
\url{experimental/outlier_migrate/phase0/preregister_om_phase0.md}. The model
is \texttt{ibm-granite/granite-4.0-h-tiny} at HuggingFace snapshot commit
\texttt{791e0d3d28c86e106c9b6e0b4cecdee0375b6124}. The prompt set contains 12
deterministically selected AIME-2025 traces, indices 0--11, with prompt SHA
\texttt{sha256:2f27c54baa8448e033d6e82f53f775dc6abe38188e4f1e5c0b97e3c74fe7c1dd}
from
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/prompt_manifest.json}.

Phase 1 follows the frozen preregistration at
\url{experimental/outlier_migrate/phase1/preregister_om_phase1.md}. The model
is \texttt{ibm-granite/granite-4.0-h-small} at HuggingFace snapshot commit
\texttt{b8c0982bab7fde4eb48110f5a069527c008fab39}. The prompt set contains 24
deterministically selected AIME-2025 traces, indices 0--23, with prompt SHA
\texttt{sha256:aa038b29332b6d137d558205ee441163e7ea4cb3cc323eb705a2f5928fd2fe4e}
from
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/prompt_manifest.json}.

For every transformer layer, the runner records per-channel activation
magnitudes. Phase 0 records decode positions \{100, 500, 1000, 5000, 10000\};
Phase 1 records \{100, 500, 1000, 5000, 10000, 20000\}. Each gate selects the
top 1\% channels by mean magnitude at position 100 and computes the fraction
whose rank at the final preregistered position differs by more than two
channel positions. The reported gate metric is the trace-bootstrap aggregate
over layers with equal layer weighting.

\section{Experiments}
\label{sec:experiments}

The Phase 0 result packet is
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z}.
The checker output is
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/checker_result.json},
the metrics file is
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/metrics.json},
and artifact completeness is recorded in
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/artifact_check.json}.
The activation artifact SHA is
\texttt{sha256:2783aa329d82e8f43ea4f342e52f0a25fc283b0b8971880f2da093739474d687}
from
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/activation_magnitude_manifest.json}.

The Phase 1 result packet is
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z}.
The checker output is
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/checker_result.json},
the metrics file is
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/metrics.json},
and artifact completeness is recorded in
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/artifact_check.json}.
The activation artifact SHA is
\texttt{sha256:326a04351e0dda28cf919fe4745d7d9341011db13d940a06cf9c8470633003e1}
from
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/activation_magnitude_manifest.json}.

The partial Phase 2 result packet is
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z}.
The checker output is
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z/checker_result.json},
the metrics file is
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z/metrics.json},
and artifact completeness is recorded in
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z/artifact_check.json}.
The model is
\url{nvidia/NVIDIA-Nemotron-3-Nano-30B-A3B-BF16} at snapshot commit
\texttt{cbd3fa9f933d55ef16a84236559f4ee2a0526848}. The activation artifact
SHA is
\texttt{sha256:bc841b938a125f05649f84577c0ddb7ab287213a683892cb81ea6ca06296206d}
from
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z/activation_magnitude_manifest.json}.
This packet is explicitly scoped as partial Nemotron-3-only validation:
Qwen3.6 and Kimi Linear are deferred, and
\texttt{full\_cross\_validation\_complete=false}.

The preregistered dynamic-outlier pass rule in both gates requires migration
fraction at least 0.05 and bootstrap CI lower bound greater than 0.05. Phase 1
uses the same dynamic path because Phase 0 returned
\texttt{PASS\_OM\_PHASE0\_DECODE\_TIME\_MIGRATION}.

\section{Results}

\begin{table}[t]
\centering
\small
\begin{tabular}{>{\raggedright\arraybackslash}p{0.24\linewidth}
                >{\raggedright\arraybackslash}p{0.32\linewidth}
                >{\raggedright\arraybackslash}p{0.32\linewidth}}
\toprule
Quantity & Phase 0 & Phase 1 \\
\midrule
Decision & dynamic-outlier pass & replicated-at-scale pass \\
Strict set-leaving fraction & 0.634244791667 & 0.566234756098 \\
Strict set-leaving 95\% CI & [0.605208333333, 0.664192708333] &
[0.550076219512, 0.581707317073] \\
Within-set rank-shuffling fraction & 0.175260416667 & 0.270934959350 \\
Within-set rank-shuffling 95\% CI & [0.160156250000, 0.191015625000] &
[0.260797764228, 0.280614837398] \\
Original preregistered migration fraction & 0.8178385416666667 & 0.843165650406504 \\
Original migration 95\% CI & [0.797265625, 0.8368489583333334] &
[0.8334349593495936, 0.8511432926829268] \\
Preregistered threshold & $\geq$0.05 and CI lower $>$0.05 &
$\geq$0.05 and CI lower $>$0.05 \\
Artifact complete & true & true \\
Model & \url{ibm-granite/granite-4.0-h-tiny} &
\url{ibm-granite/granite-4.0-h-small} \\
Trace count & 12 & 24 \\
Layer count / hidden size & 40 / 1536 & 40 / 4096 \\
Decode positions & 100, 500, 1000, 5000, 10000 &
100, 500, 1000, 5000, 10000, 20000 \\
Bootstrap seed & 20260508 & 20260508 \\
\bottomrule
\end{tabular}
\caption{OutlierMigrate Phase 0 and Phase 1 gate results. Phase 0 values are
from
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/checker_result.json}
and
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/metrics.json};
strict decomposition values are from
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/migration_decomposition.md}.
Phase 1 values are from
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/checker_result.json}
and
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/metrics.json};
strict decomposition values are from
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/migration_decomposition.md}.}
\label{tab:phase-results}
\end{table}

The original preregistered migration fraction intentionally conflates two
effects: channels can leave the top-1\% set entirely, or they can remain in
the set while shuffling rank by more than two positions. For static channel
protection, the stricter set-membership readout is the more direct
post-hoc interpretability measure, not the preregistered gate criterion.
A channel is counted as strict set-leaving only if it is in the
per-trace, per-layer top-1\% set at decode position 100 and has final rank
greater than the top-1\% boundary at decode position 10000 in Phase 0 or
20000 in Phase 1. Under that definition, 0.634244791667 of Phase 0 base
top-1\% channels and 0.566234756098 of Phase 1 base top-1\% channels leave
the protected set entirely. The within-set shuffling rates are
0.175260416667 and 0.270934959350, respectively. The gate decision remains
based only on the original preregistered migration metric.

The exact Phase 0 provenance values are: model snapshot commit
\url{791e0d3d28c86e106c9b6e0b4cecdee0375b6124}, prompt SHA
\url{sha256:2f27c54baa8448e033d6e82f53f775dc6abe38188e4f1e5c0b97e3c74fe7c1dd},
and activation artifact SHA
\url{sha256:2783aa329d82e8f43ea4f342e52f0a25fc283b0b8971880f2da093739474d687};
these are recorded in
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/model_provenance.json},
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/prompt_manifest.json},
and
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z/activation_magnitude_manifest.json}.
The exact Phase 1 provenance values are: model snapshot commit
\url{b8c0982bab7fde4eb48110f5a069527c008fab39}, prompt SHA
\url{sha256:aa038b29332b6d137d558205ee441163e7ea4cb3cc323eb705a2f5928fd2fe4e},
and activation artifact SHA
\url{sha256:326a04351e0dda28cf919fe4745d7d9341011db13d940a06cf9c8470633003e1};
these are recorded in
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/model_provenance.json},
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/prompt_manifest.json},
and
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z/activation_magnitude_manifest.json}.

The Phase 0 checker returned
\texttt{PASS\_OM\_PHASE0\_DECODE\_TIME\_MIGRATION}. Its exact artifact reason
was that the point estimate 0.81783854 is at least 0.05 and the CI lower bound
0.79726562 is greater than 0.05. The Phase 1 checker returned
\texttt{PASS\_OM\_PHASE1\_REPLICATED\_AT\_SCALE}. Its exact artifact reason
was that the point estimate 0.84316565 is at least 0.05 and the CI lower bound
0.83343496 is greater than 0.05.

\begin{table}[t]
\centering
\small
\begin{tabular}{>{\raggedright\arraybackslash}p{0.42\linewidth}
                >{\raggedright\arraybackslash}p{0.43\linewidth}}
\toprule
Quantity & Partial Phase 2 Nemotron-3 \\
\midrule
Decision & partial PASS; Qwen3.6/Kimi deferred \\
Strict set-leaving fraction & 0.533713200380 \\
Strict set-leaving 95\% CI & [0.467414529915, 0.599982193732] \\
Within-set rank-shuffling fraction & 0.269082383666 \\
Within-set rank-shuffling 95\% CI & [0.238841405508, 0.299056267806] \\
Original preregistered migration fraction & 0.820809591642925 \\
Original migration 95\% CI & [0.7865325261158594, 0.8544931149097815] \\
Preregistered threshold & $\geq$0.05 and CI lower $>$0.05 \\
Artifact complete & true \\
Model & \url{nvidia/NVIDIA-Nemotron-3-Nano-30B-A3B-BF16} \\
Trace count & 24 \\
Layer count / hidden size & 52 / 2688 \\
Decode positions & 100, 500, 1000, 5000, 10000, 20000 \\
\bottomrule
\end{tabular}
\caption{Partial Phase 2 cross-family validation. The checker-backed packet is
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z}.
This is not full cross-validation because Qwen3.6 and Kimi Linear are deferred.}
\label{tab:phase2-partial}
\end{table}

\section{Discussion}

On this Granite rank-migration surface, the Phase 0 result weakens a fixed
position-100 outlier-map interpretation: under the preregistered metric, most
position-100 top-1\% channels do not remain rank-stationary through position
10000. Phase 1 promotes the dynamic-outlier hypothesis within the Granite
family because the effect direction replicated on a larger model, twice as
many traces, and an extended decode-position grid. This is not a
first-observation claim about Mamba broadly. QMamba and OuroMamba already
establish dynamic hidden-state or activation-outlier behavior in vision Mamba.
The new evidence here is that the dynamic regime also appears under a frozen
metric in hybrid LLM long-reasoning traces.

This is the strongest OutlierMigrate measurement evidence so far, but it only
supports an observational paper candidate. The Nemotron-3 packet reduces the
risk that the signal is merely a Granite-family artifact, but it remains
partial: Qwen3.6 and Kimi Linear validation are deferred, and a
migration-aware intervention still has not been shown to improve model quality,
latency, memory, or robustness. Those two gaps are blockers for reading this
as a positive-method or systems paper. The highest-priority next gate is either
completing the deferred cross-validation under a compatible runtime or turning
the measured migration into a predeclared protection/update method.

The theoretical mechanism is plausible but not yet empirically validated
outside Granite. In channel-wise gated recurrence or linear-attention layers,
independent per-channel decay rates imply that channels retain state at
different rates. Low-decay channels can accumulate magnitude over long decode,
while high-decay channels shed magnitude. Under heterogeneous inputs and
heterogeneous decay rates, the composition of the top 1\% channels by
magnitude should therefore change across decode positions rather than remain
fixed to the position-100 set. This mechanism makes dynamic outlier migration
a natural concern for static channel-protection schemes: a calibration-time
set can be correct early in a trace and stale later in the same trace. Kimi
Linear's KDA, Qwen3.6-style Gated DeltaNet blocks, and GLA-style gated
recurrence provide architectural examples
where channel-wise or fine-grained data-dependent gating is a first-class
operation \citep{kimilinear2025,qwen36hf,yang2024gdn,yang2023gla}. This
paragraph is a mechanism argument only; this draft has not measured migration
on Kimi Linear, Qwen3.6, or GLA variants.

\section{Limitations}

This draft is not camera-ready final and should not be read as a positive
systems method. It reports two passed observational measurement gates in one
model family and one partial cross-family validation on Nemotron-3. It still
lacks Qwen3.6 and Kimi Linear validation, contamination audit, independent seed
repeat beyond the recorded bootstrap procedure, and any intervention showing
that migration-aware handling improves an application metric. The missing
intervention and deferred Qwen3.6/Kimi validation are explicit submission
blockers for a positive-method claim. The result may also depend on the chosen
decode positions, top-channel fraction, and rank-delta threshold, all of which
were fixed before analysis and should not be retuned on these packets.

The primary empirical measurement here is on Granite Mamba-2-style SSM
hybrids. Nemotron-3 provides one additional hybrid family check, but it is not
a complete substitute for validation on delta-rule linear-attention hybrids
such as Qwen3.6 and Kimi Linear. Generalization to RWKV-7, GLA-style systems,
and other efficient-recurrence variants is not established by these results.

\section{Reproducibility Statement}

The Phase 0 result packet is
\url{experimental/outlier_migrate/phase0/results/om_phase0_20260508T011824Z}.
It includes \texttt{environment.json}, \texttt{model\_provenance.json},
\texttt{prompt\_manifest.json}, \texttt{command\_metadata.json},
\texttt{random\_seed.json}, \texttt{metrics.json},
\texttt{bootstrap\_ci.json}, \texttt{checker\_result.json}, and
\texttt{artifact\_check.json}. The recorded Phase 0 command was:

\begin{verbatim}
experimental/shared/run_phase0_branch.py --branch outlier_migrate
\end{verbatim}

The Phase 1 result packet is
\url{experimental/outlier_migrate/phase1/results/om_phase1_20260508T014959Z}.
It includes the same artifact classes and records the Granite-4.0-H-Small
snapshot commit
\texttt{b8c0982bab7fde4eb48110f5a069527c008fab39}. The recorded Phase 1
command was:

\begin{verbatim}
experimental/outlier_migrate/phase1/run_om_phase1.py
\end{verbatim}

The checker decisions and exact numbers in this draft are taken from the two
Granite result directories above and the partial Phase 2 directory
\url{experimental/outlier_migrate/phase2/results/om_phase2_nemotron3_20260508T231723Z}.
The OutlierMigrate runners emit the artifact-checked metric file
\texttt{metrics.json}; they do not emit a separate
\texttt{profiler\_metrics.json}.
The recorded Phase 2 command was:

\begin{verbatim}
experimental/outlier_migrate/phase2/run_cross_model.py \
  --partial-nemotron3-only \
  --run-id om_phase2_nemotron3_20260508T231723Z \
  --batch-size 4 --dtype bfloat16 --device auto
\end{verbatim}

\section{Ethics Statement}

This is an activation-analysis measurement study on benchmark prompts. The
current artifact does not make deployment, user-data, safety, systems-efficiency,
or capability-improvement claims. The main ethical risk is overclaiming
same-family Phase 0 and Phase 1 evidence, plus the partial Nemotron-3 packet,
as a validated cross-model positive method; this draft explicitly rejects that
interpretation until deferred validation, audits, and an intervention pass.

\begin{thebibliography}{16}

\bibitem[Li et~al.(2025)]{li2025qmamba}
Yinglong Li, Xiaoyu Liu, Jiacheng Li, Ruikang Xu, Yinda Chen, and Zhiwei
Xiong.
\newblock QMamba: Post-Training Quantization for Vision State Space Models.
\newblock arXiv:2501.13624, January 2025.
\newblock \url{https://arxiv.org/abs/2501.13624}.

\bibitem[Ramachandran et~al.(2025)]{ramachandran2025ouromamba}
Akshat Ramachandran, Mingyu Lee, Huan Xu, Souvik Kundu, and Tushar Krishna.
\newblock OuroMamba: A Data-Free Quantization Framework for Vision Mamba.
\newblock arXiv:2503.10959, ICCV 2025.
\newblock \url{https://arxiv.org/abs/2503.10959}.

\bibitem[Pierro and Abreu(2024)]{pierro2024mambaptq}
Alessandro Pierro and Steven Abreu.
\newblock Mamba-PTQ: Outlier Channels in Recurrent Large Language Models.
\newblock arXiv:2407.12397, 2024.
\newblock \url{https://arxiv.org/abs/2407.12397}.

\bibitem[Chiang et~al.(2024)]{chiang2024quamba}
Hung-Yueh Chiang, Chi-Chih Chang, Natalia Frumkin, Kai-Chiang Wu, and Diana
Marculescu.
\newblock Quamba: A Post-Training Quantization Recipe for Selective State
Space Models.
\newblock arXiv:2410.13229, 2024.
\newblock \url{https://arxiv.org/abs/2410.13229}.

\bibitem[Xu et~al.(2025)]{xu2025mambaquant}
Zukang Xu, Yuxuan Yue, Xing Hu, Zhihang Yuan, Zixu Jiang, Zhixuan Chen,
Jiangyong Yu, Chen Xu, Sifan Zhou, and Dawei Yang.
\newblock MambaQuant: Quantizing the Mamba Family with Variance Aligned
Rotation Methods.
\newblock arXiv:2501.13484, 2025.
\newblock \url{https://arxiv.org/abs/2501.13484}.

\bibitem[Chen and Hemani(2026)]{chen2026quambase}
Yizhi Chen and Ahmed Hemani.
\newblock Late Breaking Results: Quamba-SE: Soft-edge Quantizer for
Activations in State Space Models.
\newblock arXiv:2601.09451, DATE Late Breaking Results 2026.
\newblock \url{https://arxiv.org/abs/2601.09451}.

\bibitem[Xiao et~al.(2023)]{xiao2023smoothquant}
Guangxuan Xiao, Ji Lin, Mickael Seznec, Hao Wu, Julien Demouth, and Song Han.
\newblock SmoothQuant: Accurate and Efficient Post-Training Quantization for
Large Language Models.
\newblock arXiv:2211.10438, ICML 2023.
\newblock \url{https://arxiv.org/abs/2211.10438}.

\bibitem[Lin et~al.(2023)]{lin2023awq}
Ji Lin, Jiaming Tang, Haotian Tang, Shang Yang, Wei-Ming Chen, Wei-Chen Wang,
Guangxuan Xiao, Xingyu Dang, Chuang Gan, and Song Han.
\newblock AWQ: Activation-aware Weight Quantization for LLM Compression and
Acceleration.
\newblock arXiv:2306.00978, MLSys 2024.
\newblock \url{https://arxiv.org/abs/2306.00978}.

\bibitem[Ashkboos et~al.(2024)]{ashkboos2024quarot}
Saleh Ashkboos, Amirkeivan Mohtashami, Maximilian L. Croci, Bo Li, Pashmina
Cameron, Martin Jaggi, Dan Alistarh, Torsten Hoefler, and James Hensman.
\newblock QuaRot: Outlier-Free 4-Bit Inference in Rotated LLMs.
\newblock arXiv:2404.00456, 2024.
\newblock \url{https://arxiv.org/abs/2404.00456}.

\bibitem[Hooper et~al.(2024)]{hooper2024kvquant}
Coleman Hooper, Sehoon Kim, Hiva Mohammadzadeh, Michael W. Mahoney, Yakun
Sophia Shao, Kurt Keutzer, and Amir Gholami.
\newblock KVQuant: Towards 10 Million Context Length LLM Inference with KV
Cache Quantization.
\newblock arXiv:2401.18079, NeurIPS 2024.
\newblock \url{https://arxiv.org/abs/2401.18079}.

\bibitem[Jang and Tambe(2025)]{jang2025blockdialect}
Wonsuk Jang and Thierry Tambe.
\newblock BlockDialect: Block-wise Fine-grained Mixed Format Quantization for
Energy-Efficient LLM Inference.
\newblock arXiv:2501.01144, ICML 2025.
\newblock \url{https://arxiv.org/abs/2501.01144}.

\bibitem[Qwen Team(2026)]{qwen36hf}
Qwen Team.
\newblock Qwen3.6-35B-A3B model card.
\newblock Hugging Face, 2026.
\newblock \url{https://huggingface.co/Qwen/Qwen3.6-35B-A3B}.

\bibitem[Yang et~al.(2023)]{yang2023gla}
Songlin Yang, Bailin Wang, Yikang Shen, Rameswar Panda, and Yoon Kim.
\newblock Gated Linear Attention Transformers with Hardware-Efficient
Training.
\newblock arXiv:2312.06635, 2023.
\newblock \url{https://arxiv.org/abs/2312.06635}.

\bibitem[Yang et~al.(2024)]{yang2024gdn}
Songlin Yang, Jan Kautz, and Ali Hatamizadeh.
\newblock Gated Delta Networks: Improving Mamba2 with Delta Rule.
\newblock arXiv:2412.06464, 2024.
\newblock \url{https://arxiv.org/abs/2412.06464}.

\bibitem[Moonshot AI Kimi Team(2025)]{kimilinear2025}
Moonshot AI Kimi Team: Yu Zhang et~al.
\newblock Kimi Linear: An Expressive, Efficient Attention Architecture.
\newblock arXiv:2510.26692, October 2025.
\newblock \url{https://arxiv.org/abs/2510.26692}.

\end{thebibliography}

\end{document}
