\documentclass{article}

\usepackage[submission]{../../../colm_final/paper/colm2026_conference}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{url}

\title{HORN: Directional Outlier Propagation in Hybrid Reasoners}
\author{Anonymous authors\\Paper under double-blind review}

\begin{document}
\maketitle

\begin{abstract}
This is a living paper shell for a preregistered Mac-first experiment. HORN
tests whether attention-to-SSM and SSM-to-attention boundaries differ in
outlier magnitude and quantization-noise sensitivity. No positive result or
native systems claim is made yet.
\end{abstract}

\section{Current Gate}

The branch must pass activation-magnitude, quantization-noise propagation, and
cross-model/control gates before GPU validation.
H1a first screens one live hybrid model for directional activation statistics,
H1 requires the same selected direction on at least two hybrid models, H2 tests matched
FP4-equivalent noise propagation, and H3 requires the direction to survive
hybrid-model repeats while weakening on pure-attention and pure-Mamba controls.

\section{Mac Artifact Status}

A deterministic synthetic H1a packet validates the directional-boundary
real-schema contract. It emits 72 rows, selected max-abs ratio 4.044,
non-boundary selected-direction control ratio 1.042, and permuted
selected-direction control ratio 0.247, producing
\texttt{SCHEMA\_REHEARSAL\_NOT\_PROMOTABLE\_SYNTHETIC\_HORN\_H1A}. This is
synthetic-only: it is not model evidence, not a quantization-noise result, and
not GPU evidence.
The metadata-only model eligibility packet identifies live hybrid targets and
records architecture-map hashes for admissible packets, but no target weights
are cached repo-locally, so HORN has not yet produced real boundary rows on the
current Mac.

\section{Next Evidence}

The next admissible row is a real H1a boundary activation dump from a hybrid model.
HORN remains a control branch unless real H1/H2/H3 gates show consistent
directional asymmetry that pure-attention and pure-Mamba controls do not share.
Pure-architecture controls are an H3 requirement, not a blocker for the first
H1a boundary-activation packet.

\section{Admissible Real Packet}

The real H1a/H1 packet must use the shared config-derived architecture maps rather
than substring-only module classification. Each row must report
\texttt{model\_id}, \texttt{prompt\_id}, adjacent layer IDs, \texttt{direction},
\texttt{matched\_boundary\_direction}, \texttt{boundary\_index}, normalization placement, \texttt{max\_abs},
\texttt{rms}, \texttt{kurtosis}, and \texttt{control\_type}. The required
controls are boundary rows, non-boundary adjacent rows, and permuted-direction
rows. A valid packet must cover both \texttt{attention->ssm} and
\texttt{ssm->attention} in boundary rows and in matched control labels.
Non-boundary controls may retain their true architecture direction, such as
\texttt{ssm->ssm}, while \texttt{matched\_boundary\_direction} records the
boundary direction they control for, and both matched directions must appear
for every prompt; each
permuted-direction row must match an observed boundary tuple with the same
prompt ID, layer IDs, normalization positions, and measured metrics while
flipping the actual \texttt{direction} label only. It cannot
pass if the permuted controls preserve the selected high-magnitude direction;
a faithful label flip may preserve unsigned asymmetry while moving the signal
to the opposite label. Non-boundary controls must stay below the selected H1
threshold, not merely below the observed boundary ratio. It must also
include at least 12 fixed prompt IDs unless
the packet records an explicit resource limit. Any resource-limited packet must
use a \texttt{RESOURCE\_LIMITED\_NOT\_PROMOTABLE} decision. The packet must
include \texttt{prompt\_ids\_hash}, \texttt{architecture\_map\_hash},
\texttt{trace\_plan\_hash}, and a
summary with prompt count, boundary directions, selected H1a ratio, selected H1a
CI lower bound, support fraction, non-boundary control ratio, and
permuted-direction ratio, then pass
\texttt{experimental.shared.check\_gate\_packet --mode real --project horn}.
The checker recomputes these H1a fields from raw rows and rejects stale or
hand-filled summaries.
Saved tensor packets should be converted with
\texttt{experimental.shared.hybrid\_trace\_packet\_builder} before validation.
The current capture-manifest templates are generated by
\texttt{experimental.shared.hybrid\_trace\_capture\_manifest}; they must be
filled from real boundary-activation captures before packet building. Served HF
IDs are preserved as \texttt{served\_model\_id} while rows are validated
against the canonical architecture-map \texttt{model\_id}; permuted-direction
entries use \texttt{tensor\_alias\_of} to reuse the observed boundary tensor.

\begin{table}[h]
\centering
\small
\begin{tabular}{ll}
\toprule
Required control & Reviewer objection addressed \\
\midrule
Explicit boundary map & Avoids substring artifacts. \\
Non-boundary adjacent pair & Separates boundary effects from depth effects. \\
Permuted direction label & Tests whether direction carries signal. \\
Matched normalization side & Controls RMSNorm/LayerNorm placement. \\
Pure-architecture control & Prevents generic outlier claims. \\
\bottomrule
\end{tabular}
\caption{Controls required before HORN can claim directional hybrid asymmetry.}
\end{table}

\section{Reviewer Packet}

The reviewer-facing claim boundary is tracked in
\url{experimental/horn/paper/reviewer_pack.md}. It records that HORN is not camera-ready as a method or measurement paper until real H1--H3 evidence exists.

\end{document}
