\documentclass{article}

\usepackage[submission]{../../../colm_final/paper/colm2026_conference}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{url}

\title{SSQ-LR: Sub-FP16 Recurrent State Quantization for Hybrid Reasoners}
\author{Anonymous authors\\Paper under double-blind review}

\begin{document}
\maketitle

\begin{abstract}
This is a living paper shell for a preregistered Mac-first experiment. SSQ-LR
tests whether recurrent SSM state in hybrid reasoning models can be quantized
below FP16 while preserving quality. No positive result or native systems claim
is made yet; current S2/S3 scouts stop the frozen recipe, and no recipe clears
quality, byte accounting, and no-retuning transfer.
\end{abstract}

\section{Current Gate and Claim Boundary}

The branch must pass state-distribution heterogeneity, state-only quantization
sensitivity, and cross-model transfer gates before any GPU validation.
S1 must first show real state heterogeneity on a live hybrid model; S2 then
freezes one state-only quantization recipe; S3 tests that recipe without
per-model retuning. Failing any gate kills the branch before GPU work.
The current real trace packet builder/checker covers S1. S2 and S3 now have
executable follow-up contracts in \texttt{experimental/shared/followup\_gate\_contracts.py}:
they require frozen recipe IDs, byte accounting, BF16 no-op controls, paired
accuracy and continuation-NLL quality bounds, verbosity/length-drift readouts
before revival, frozen recipe/source hashes, and no-retuning transfer rows.
The byte gate is counted as
\[
\frac{\mathrm{BF16\ state\ bytes}}{\mathrm{quantized\ state\ bytes} + \mathrm{scale\ bytes} + \mathrm{metadata\ bytes}},
\]
so scale tensors, layer masks, and recipe metadata are not free.
Resource-limited S2 scouts are not current evidence
for promotion. SSQ-LR makes no quality, byte-savings, transfer, GPU, latency, or
throughput claim until non-resource-limited S1--S3 packets clear.

\section{Mac Artifact Status}

A deterministic synthetic S1 packet now validates the real-schema checker path
before real hybrid-model state dumps. It emits 288 rows covering 12 prompts, 6
recurrent layers, and the four preregistered buckets, producing
\texttt{SCHEMA\_REHEARSAL\_NOT\_PROMOTABLE\_SYNTHETIC\_SSQ\_LR\_S1}. This is
synthetic-only: it is not model evidence, not a quality result, and not GPU
evidence.
The held-out S1b packet is the strongest diagnostic heterogeneity evidence so far: it repeats
the frozen selected layers \texttt{0}, \texttt{12}, and \texttt{30} on 12
held-out prompts with selected ratio 2.459 and lower bound 1.861. Uniform
state recipes do not yet convert that heterogeneity into a method. MXFP4 preserves short
BF16-argmax fidelity but reaches only 3.765--3.938$\times$ state-memory
reduction after scale bytes. INT3 reaches 4.923--5.224$\times$ after scale
bytes, but both 12-prompt held-out scouts fail because at least one prompt loses
argmax fidelity. The 4-prompt INT3 pass is therefore treated as resource-limited
and non-promoting.
The newest resource-limited mixed-block S2b scout selects
\texttt{mixed\_int3\_mxfp4\_low\_error\_25pct}: it reaches 4.192$\times$
counted state-memory reduction on the 12-prompt held-out replay with zero
BF16-argmax delta and 0.03956 selected NLL-delta CI high. This short-window
pass is still simulated replay and cannot promote to GPU.
The same recipe then fails a longer-window replay
(\texttt{--max-input-tokens 24 --prefix-tokens 8}): it preserves the
4.192$\times$ byte readout but has selected accuracy CI high 0.0667 and
selected NLL-delta CI high 0.0764. The mixed recipe is therefore stopped before
GPU validation.
Layer-localization gives one narrower layer set: layers 0 and 30 individually
pass the same longer-window replay with INT3 state at 5.224$\times$ counted
memory reduction and zero selected accuracy delta, while layer 12 fails. The
combined layers 0 and 30 replay also passes with selected NLL-delta CI high
0.04294. A stricter prefilter replay on this layer set weakens pure INT3
(accuracy CI high 0.105) but selects
\texttt{mixed\_int3\_mxfp4\_low\_error\_25pct} at 4.192$\times$ counted
memory, zero selected accuracy drift, and 0.05044 selected NLL-delta CI high.
The S3 transfer replay then fails on the second complete local hybrid model:
Granite 350M with the same 12 prompts, layers 0 and 30, 32-token context, and
no retuning selects an INT8 fallback at only 1.984$\times$ memory. Layer
localization is post-hoc diagnostic evidence only: layer 0 can pass on Granite
350M and layer 30 fails, but the source Granite Tiny layer-0 replay and Granite
350M layer-0 replay prefer different low-bit recipes. Two local S3 packets that
freeze either mixed25 or INT3 for layer 0 both fail, so there is no current
frozen recipe that transfers.

\begin{table}[h]
\centering
\small
\begin{tabular}{p{0.34\linewidth}p{0.23\linewidth}rr}
\toprule
Mac packet & Selected recipe & Memory & Accuracy CI high \\
\midrule
Granite Tiny layer 0 source & mixed25 INT3/MXFP4 & 4.192$\times$ & 0.000 \\
Granite 350M layers 0,30 & INT8 fallback & 1.984$\times$ & 0.000 \\
Granite 350M layer 0 & INT3 & 5.224$\times$ & 0.000 \\
Granite 350M layer 30 & INT8 fallback & 1.984$\times$ & 0.000 \\
Two-model S3, frozen mixed25 layer 0 & mixed25 INT3/MXFP4 & -- & 0.0667 \\
Two-model S3, frozen INT3 layer 0 & INT3 & -- & 0.0526 \\
\bottomrule
\end{tabular}
\caption{Current SSQ-LR transfer failure. Layer diagnostics are not a rescue:
the recipe that passes one Granite model does not pass the other under the
frozen no-retuning S3 contract. The rows are Mac-local simulated replay, not
native GPU, bandwidth, or serving evidence.}
\end{table}

\section{Next Evidence}

The next admissible action is not GPU validation. SSQ-LR needs a new
preregistered recipe or layer-selection rule that is frozen before seeing
transfer rows, then must rerun no-retuning transfer plus verbosity/length-drift
readouts. The current S1b/S2b result is diagnostic evidence, not a positive
method.

\section{Admissible Real Packet}

Promotable packets must pin model/tokenizer revisions, prompt and trace-plan
hashes, the shared architecture-map hash, exact command/device/dtype metadata,
and saved recurrent-state tensor hashes. S1 rows must cover all four registered
position buckets, including \texttt{prefill\_end}, for every prompt/layer pair
and at least 12 prompt IDs unless explicitly marked resource-limited. The
checker recomputes row metrics and gate aggregates from raw tensors, rejects
stale summaries, and requires
\texttt{RESOURCE\_LIMITED\_NOT\_PROMOTABLE} decisions for any subset packet.
Full schema details live in the README and runbook.

\begin{table}[h]
\centering
\small
\begin{tabular}{p{0.34\linewidth}p{0.56\linewidth}}
\toprule
Required baseline & Purpose \\
\midrule
BF16 no-op replay & Catch hook/serialization bugs. \\
FP16 stochastic-rounding SSM cache & Deployment baseline from Nemotron-style hybrid serving. \\
INT16 block-scaled SSM cache & Higher-precision recurrent-cache control. \\
INT8 state-only & Standard low-bit state control. \\
FP8-style state-only & Checks whether sub-FP16 state can match the likely deployed fallback. \\
MXFP4 state-only & Candidate sub-FP16 recipe. \\
INT3 state-only & Tests whether sub-4-bit state can clear scale-byte overhead. \\
Random same-L2 noise & Separate structure from noise magnitude. \\
Scale shuffle & Test scale assignment sensitivity. \\
Byte accounting & Include scale/metadata overhead. \\
\bottomrule
\end{tabular}
\caption{Controls required before SSQ-LR can claim a state-quantization recipe.}
\end{table}

\section{Reviewer Packet}

The reviewer-facing claim boundary is tracked in
\url{experimental/ssq_lr/paper/reviewer_pack.md}. It records that SSQ-LR is
not camera-ready as a method paper until non-resource-limited, passing S1--S3 evidence exists.
The newest novelty baseline is Nemotron 3 Super's recurrent-cache recipe:
NVIDIA documentation treats Mamba SSM cache quantization as a separate
deployment challenge and selects FP16 with stochastic rounding for the SSM cache,
with INT16/block-scaled variants discussed as controls. SSQ-LR is therefore not
novel for noticing SSM cache precision; it must beat this baseline with a
sub-FP16 recipe and paired quality evidence. The current local recipes do not
beat this baseline. The source memo is
\url{../../../references/754_ssq_lr_nemotron_state_cache_refs_20260507.md}.

\section{Claim Boundary}

SSQ-LR is only about recurrent SSM-state precision in hybrid reasoners. It
does not claim weight quantization, activation quantization, KV-cache
quantization, native serving speed, or memory savings until a real state recipe
and later GPU validation pass their preregistered gates.

\end{document}
