\documentclass{article}

\usepackage[submission]{../../../colm_final/paper/colm2026_conference}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{url}

\title{HBSM Preregistration Note: Hybrid Boundary Sensitivity Mechanisms}
\author{Anonymous authors\\Paper under double-blind review}

\begin{document}
\maketitle

\begin{abstract}
This is a living paper shell for a preregistered Mac-first experiment. HBSM
tests whether sensitivity in the planned hybrid-reasoner target family can be
explained and predicted cheaply. The broad forward-only sensitivity claim is wounded by
KL Lens-style forward-sensitivity work, so no novelty or positive result is
claimed yet. Current Granite Tiny resource-limited scouts fail B1.
\end{abstract}

\section{Current Gate}

The branch must pass sensitivity heterogeneity, no-forward predictor, and
mechanism gates. If it overlaps with HORN, it should be folded into a unified
hybrid-quantization paper.
The current broad hypothesis is demoted after failed Granite Tiny B1 scouts;
the thresholds below are retained only as future admissibility rules for a
narrower preregistered mechanism hypothesis.
B1 must first reproduce sensitivity heterogeneity on current hybrid reasoners,
B2 must beat no-forward baselines without overfitting depth or size, and B3
must add mechanism beyond HORN's boundary-direction test.
The current real trace packet builder/checker covers B1. B2 and B3 now have
executable follow-up contracts in \texttt{experimental/shared/followup\_gate\_contracts.py}:
they require predictor registry hashes, fixed hyperparameter hashes, train-only
selection, held-out Spearman and baseline-margin fields, HORN-alignment signs,
noise-off controls, and direction-flip controls. These contracts are not model
evidence: no cheap-predictor, no-forward-pass, or mechanistic claim is allowed
until real B1 promotes and B2/B3 are filled with train/test or
leave-one-model-out predictor splits and matched-noise attention-versus-SSM
controls.

\section{Mac Artifact Status}

A deterministic synthetic B1 real-schema rehearsal now validates the
layer-sensitivity packet contract. It emits 720 rows, including 480 primary
prompt rows and 240 layer-aligned control rows, and aggregates those primary rows to 40
scoring layers. The packet passes the real checker in non-promoting mode with
\texttt{SCHEMA\_REHEARSAL\_NOT\_PROMOTABLE\_SYNTHETIC\_HBSM\_B1}. This is
synthetic-only: it is not model evidence, not a novelty claim, and not GPU
evidence.
The evaluator derives measured top-decile membership from aggregated
\texttt{kl\_or\_nll\_drift}; supplied \texttt{top\_decile\_flag} values are
accepted only when they match that measured ranking for every prompt row inside
each measured layer.
The metadata-only model eligibility packet identifies the current live hybrid
targets and architecture hashes, but no corresponding weights are cached
repo-locally except Granite Tiny. HBSM now has two Granite Tiny
resource-limited forward-sensitivity packets: a one-prompt smoke and a
two-prompt prompt-repeat scout. Both are checker-valid, non-promoting B1
failures; the 8-layer smoke is too small to formally satisfy the full Fisher
gate, but the observed rank direction and cheap-predictor sign are wrong. There
is still no real forward-sensitivity evidence that promotes B1. The two-prompt
scout removes the one-prompt boundary-top-decile hint:
its boundary top-decile count is zero, Fisher \(p=1.0\), and cheap-predictor
Spearman is \(-0.667\).

\section{Next Evidence}

The next admissible HBSM action is not GPU validation. Either preregister a
narrower mechanism hypothesis that explains why the two-prompt B1 scout failed,
or fold HBSM into the negative/control appendix of the hybrid-quantization
workstream.

\section{Admissible Real Packet}

The real B1 packet must use fixed boundary flags from the shared architecture
maps. Each row must report \texttt{model\_id}, \texttt{layer},
\texttt{boundary\_flag}, \texttt{precision\_perturbation},
\texttt{kl\_or\_nll\_drift}, \texttt{cheap\_predictor},
\texttt{parameter\_count}, \texttt{weight\_norm},
\texttt{top\_decile\_flag}, \texttt{random\_top\_decile},
\texttt{train\_test\_split}, and \texttt{control\_type}.
The packet must include \texttt{config.json}, \texttt{raw\_rows.jsonl},
\texttt{summary.json}, \texttt{summary.md}, \texttt{decision.md},
\texttt{prompt\_ids\_hash}, \texttt{architecture\_map\_hash},
\texttt{trace\_plan\_hash}, the exact \texttt{trace\_plan\_path} whose file
SHA-256 matches that hash, matching \texttt{row\_count}, and pass
\texttt{experimental.shared.check\_gate\_packet --mode real --project hbsm}.
The current capture-manifest templates are generated by
\texttt{experimental.shared.hybrid\_trace\_capture\_manifest}; they must be
filled from real forward-sensitivity captures before packet building. Served HF
IDs are preserved as \texttt{served\_model\_id} while rows are validated
against the canonical architecture-map \texttt{model\_id}; unfilled template
markers are rejected before metric coercion.
The checker requires both true and false boundary flags, finite sensitivity and
predictor fields, all listed controls on the same \texttt{(model\_id, layer)}
scoring set as the primary rows, and a perturbation-off row whose drift is
near zero. It also requires exact measured top-decile cardinality
\(\lceil 0.10\cdot\mathrm{scoring\ layers}\rceil\), where scoring aggregates
primary prompt rows to one row per \texttt{(model\_id, layer)}, agreement
between supplied top-decile labels and measured drift ranking on every prompt
row, an equally sized
random baseline that is not enriched, KL-style and activation/outlier comparator
controls, and both train and test split rows unless the packet records an
explicit resource limit.
Any resource-limited packet must use a
\texttt{RESOURCE\_LIMITED\_NOT\_PROMOTABLE} decision. The summary must include
primary-row, scoring-layer, and prompt counts; top-decile count; random
top-decile count; train/test counts; boundary top-decile enrichment; Fisher
p-value; cheap-predictor Spearman correlation; baseline Spearman values; and
the cheap-predictor margin over the best baseline. The checker recomputes these
fields from \texttt{raw\_rows.jsonl} and rejects stale or hand-filled
\texttt{summary.json} entries.

\begin{table}[h]
\centering
\small
\begin{tabular}{p{0.24\linewidth}p{0.66\linewidth}}
\toprule
Gate & Promotion threshold \\
\midrule
B1 sensitivity & Primary boundary-only rows show top-decile enrichment over non-boundary rows with Fisher $p<0.05$, while matched random flags do not reproduce the enrichment. \\
B2 predictor & Spearman $\rho \ge 0.6$ with bootstrap lower bound $\ge 0.3$. \\
B3 mechanism & Matched-noise attention drift exceeds SSM drift and aligns with HORN. \\
\bottomrule
\end{tabular}
\caption{Pre-registered HBSM gate thresholds before any standalone claim.}
\end{table}

\begin{table}[h]
\centering
\small
\begin{tabular}{ll}
\toprule
Required baseline & Purpose \\
\midrule
Random top-decile flags & Tests enrichment beyond chance. \\
Layer-index baseline & Controls depth effects. \\
Parameter-count baseline & Controls layer size. \\
Weight-norm baseline & Controls simple magnitude predictors. \\
Boundary-only baseline & Tests whether HORN already explains the effect. \\
KL-style ranking baseline & Separates HBSM from forward-sensitivity ranking. \\
Activation/outlier baseline & Separates boundary sensitivity from generic outlier quantization. \\
Perturbation-off row & Catches hook and scoring bugs. \\
Train/test split & Reduces cheap-predictor overfit. \\
\bottomrule
\end{tabular}
\caption{Controls required before HBSM can claim a cheaper sensitivity predictor.}
\end{table}

\section{Reviewer Packet}

The reviewer-facing claim boundary is tracked in
\url{experimental/hbsm/paper/reviewer_pack.md}. It records that the current HBSM hypothesis is demoted; only a future narrower preregistered HBSM hypothesis with B1--B3 evidence could become a standalone paper.

\section{Novelty Boundary}

The closest risk is KL Lens-style forward sensitivity for mixed-precision
hybrid models \cite{kllens}. HBSM can survive only if it shows a cheaper
no-forward predictor or a boundary mechanism on current hybrid reasoners;
otherwise it should be folded into negative/control appendix context or killed.

\begin{thebibliography}{1}
\bibitem[KL Lens(2026)]{kllens}
Jason Kong, Nilesh Prasad Pandey, Flavio Ponzina, and Tajana Rosing.
A KL Lens on Quantization: Fast, Forward-Only Sensitivity for
Mixed-Precision SSM-Transformer Models. arXiv:2604.13440, 2026.
\url{https://arxiv.org/abs/2604.13440}.
\end{thebibliography}

\end{document}
